---
title: "DDS Analytics"
author: "Authors: Daniel Davieau, Lakeitra Webb, Emil Ramos"
date: "April 24, 2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r Environment Information}
#clear R Environment
rm(list=ls()) 

#set the working directory
setwd("~/GitHub/DDS-Analytics") 
#setwd("C:/Users/i711214/Desktop/DDS-Analytics")

#load Libraries
library(olsrr)
library(knitr)
library(readxl)
library(dygraphs)
library(tidyr)
library(dplyr)
library(ggplot2)
library(forecast)
library(xtable)
library(kableExtra)
library(readr)
library(reshape2)
library(wordcloud2)
library(investr)
library(MASS)

#Report environment Info
sessionInfo()

#establish a color palette for consistency
palette <- c('#073857','#4E91DD','#A13764','#7D9099','#32C3A4','#2F63A0','#484848','#2C707E','#D1A34F','#E34A46','#FCFCFC')
```
## Business Objective
Conduct an analysis of existing employee data to determine factors that lead to attrition. Highlight specific trends that contribute to turnover.

## Data
Data was provided by the client in form of a .csv file "CaseStudy2-data.xlsx". The data defines employee information by Education, Environment Satisfaction, Job Involvement, Job Satisfaction, Performance Rating, Relationship Satisfaction, and Work Life Balance.

####Basic Statistics
The data includes a total of 1470 current and terminated employee records with 35 variables. 
```{r}
#load the provided excel file
CaseStudy2_data <- read_excel("Input/CaseStudy2-data.xlsx")

#checking dimensions, create a variable, add columns and row labels
dims<-as.data.frame(dim(CaseStudy2_data))
cbind.data.frame("Observations",dims)
rownames(dims)<-c("Observations","Variables")
colnames(dims)<-c("Data Records")
```
```{r}
#display the data dimensions and measures
kable(dims, "html") %>%
kable_styling(bootstrap_options = "striped", full_width = F,position = "left")
```
Summary of measures included in the data:
```{r,fig.align='center'}
# Interesting variables: daily rate, education,  job satisfaction,number of companies worked, performance rating,years at company, years with current manager

#Variables to use with caution (i.e discrimination law, ethics:Age, Gender, marital status, relationship statisfaction, total working years

#create a dataframe with descriptive statistics of quantitative variables.
dr <- matrix(summary(CaseStudy2_data$DailyRate))
cw<-matrix(summary(CaseStudy2_data$NumCompaniesWorked))
yac<-matrix(summary(CaseStudy2_data$YearsAtCompany))
ywm<-matrix(summary(CaseStudy2_data$YearsWithCurrManager))
dfh<-matrix(summary(CaseStudy2_data$DistanceFromHome))
psh<-matrix(summary(CaseStudy2_data$PercentSalaryHike))
ycr<-matrix(summary(CaseStudy2_data$YearsInCurrentRole))
x<-data.frame(cbind(dr,cw,yac,ywm,dfh,psh,ycr))

#set column names
colnames(x)<-c("DailyRate","Number of Companies Worked","Years at Company","Years with Manager", "Distance From Home","Percent Salary Hike","Years in Current Role")

#set row names
rownames(x)<-c("Min.","1st Qu.","Median","Mean","3rd Qu.","Max.")
kable(x, "html") %>%
kable_styling(bootstrap_options = "striped", full_width = F)
```
##Methodology
Attrition is the central theme of this analysis. We interpreted the value "Yes" in the data provided under attrition as an indicator that the employee is terminated.

We categorized each record as "Current" or "Terminated" and look for patterns in the variables that may explain why employees become terminated.  

*__Current__ meaning currently employed by the firm.*  
*__Terminated__ meaning left the firm (voluntarily or non-voluntarily)*
```{r Add Employee Status, include=FALSE}
#Initialize Employee Status Column
EmployeeStatus<-as.character("Derp") 

#For some reason I get an error when initializing EmployeeStatus. It seems like the code is getting ahead of itself trying to cbind(CaseStudy2_data, EmployeeStatus) Until I ge ta chance to figure out how to initialize it properly I inserted this 1 second pause:
Sys.sleep(1)

#Add Employee Status to the CaseStudy2_data dataframe.
cbind(CaseStudy2_data, EmployeeStatus)

#If attrition is "Yes" then set the employee status to "Terminated"
CaseStudy2_data$EmployeeStatus[CaseStudy2_data$Attrition == "Yes"] <- "Terminated"

#If attrition is "No" then set the employee status to "Current"
CaseStudy2_data$EmployeeStatus[CaseStudy2_data$Attrition == "No"] <- "Current" 
```
```{r,fig.align='center'}
ggplot(CaseStudy2_data, aes(x=EmployeeStatus,  fill=EmployeeStatus))+
labs(title="Employees", x = "", y="Count", fill="Status") +geom_bar(stat = "count") +
  scale_fill_manual(values=c("#2F63A0","#7D9099"))
```
```{r}
class((CaseStudy2_data))
unique(CaseStudy2_data$Gender)
unique(CaseStudy2_data$YearsAtCompany)
unique(CaseStudy2_data$Attrition)
```
As requested employees under the age of 18 have been excluded from this analsyis. The table below shows the youngest age record included in the data:
```{r}
#verify no records with participants under age 18
underage<-as.matrix(min(CaseStudy2_data$Age))
colnames(underage)<-c("Youngest Age Record")
kable(underage, "html") %>%
kable_styling(bootstrap_options = "striped", full_width = F,position = "left")
```
####Distributions  
The following are the distributions of employees by various measures.
```{r,fig.align='center'}

ggplot(CaseStudy2_data, aes(x=HourlyRate,fill = EmployeeStatus))+
  labs(title="Employee Distribution by Hourly Rate", x = "Hourly Rate", y="Employee Count", fill="Status") +
geom_histogram(binwidth=7, color="white",position = position_stack(reverse = TRUE)) +
  scale_fill_manual(values=c("#2F63A0","#7D9099")) 
 
ggplot(CaseStudy2_data, aes(x=NumCompaniesWorked,fill = EmployeeStatus)) + 
  labs(title="Employee Distribution by Number of Companies Worked", x = "Number of Companies", y="Employee Count", fill="Status") +
  geom_histogram( binwidth=1, position = position_stack(reverse = TRUE),color="white") +
  scale_fill_manual(values=c("#2F63A0","#7D9099"))
```

```{r,fig.align='center'}
#Education Distribution
ggplot(CaseStudy2_data, aes(x=Education,fill = EmployeeStatus)) + 
  labs(title="Employee Distribution by Education", x = "Education Level", y="Employee Count", fill="Status") +
  geom_histogram( binwidth=1, position = position_stack(reverse = TRUE),color="white") +
  scale_fill_manual(values=c("#2F63A0","#7D9099"))
```

####Frequencies  
The following are frequencies by Gender and Job Roles
```{r,fig.align='center'}
ggplot(CaseStudy2_data, aes(x=Gender,  fill=Gender)) +geom_bar(stat = "count") +
  labs(title="Employee Frequency by Gender", x = "", y="Employee Count", fill="Status") +
  scale_fill_manual(values=c("#A13764","#2F63A0"))
```
```{r,fig.align='left'}
#Occupation Frequency
ggplot(CaseStudy2_data, aes(x=JobRole,  fill=JobRole)) +geom_bar(stat = "count") +
  labs(title="Employee Frequency by Job Role", x = "", y="Employee Count", fill="Role") +
  scale_fill_manual(values=palette) +
  coord_flip() + 
  theme(legend.position="none")
```
  
As requested we have captured the counts of management positions in the table below:
```{r}
#Need to come back and make this into a function
#Get counts of each management category

#Sales Execs
countsalesExecs<-as.matrix(length(which(CaseStudy2_data$JobRole == "Sales Executive")))
rownames(countsalesExecs)<-c("Count")
colnames(countsalesExecs)<-c("Sales Executives")

#Manufacturing Directors
countManufactureDirectors<-as.matrix(length(which(CaseStudy2_data$JobRole == "Manufacturing Director")))
rownames(countManufactureDirectors)<-c("Count")
colnames(countManufactureDirectors)<-c("Manufacturing Directors")

#Managers
countManagers<-as.matrix(length(which(CaseStudy2_data$JobRole == "Manager")))
rownames(countManagers)<-c("Count")
colnames(countManagers)<-c("Managers")

#Research Directors
countResearchDirectors<-as.matrix(length(which(CaseStudy2_data$JobRole == "Research Director")))
rownames(countResearchDirectors)<-c("Count")
colnames(countResearchDirectors)<-c("Research Directors")

#Combine the counts into a dataframe
Management<-as.data.frame(t(cbind(countsalesExecs,countManufactureDirectors,countManagers,countResearchDirectors)))
```
```{r,fig.align='left'}
#Display a table
kable(Management, "html") %>%
kable_styling(bootstrap_options = "striped", full_width = F,position = "left")
```
####Is there a relationship between Age and Income?  
There's no apparent relationship between Age and Income. The plot below shows a very slight upward incline as age increases but is relatively insignificant.  
```{r,fig.align='center'}
#referenced http://www.sthda.com/english/wiki/ggplot2-scatter-plots-quick-start-guide-r-software-and-data-visualization

#Is there a relationship between Age and Income?  Create a scatterplot and make an assessment of whether there is a relationship. Color each dot by gender
ggplot(CaseStudy2_data, aes(x=Age, y=DailyRate, color=Gender)) +
   geom_point(size=2)+
  scale_color_manual(values = c("#A13764","#2F63A0"))+
  geom_smooth(method = lm, color="#484848",lwd = 1, se=TRUE)
```



```{r,fig.align='center'}
####Is there a relationship between life satisfaction and any other attribute?  
# ggplot(CaseStudy2_data, aes(x=WorkLifeBalance,y = YearsAtCompany, color=EmployeeStatus )) +
#    geom_point(size=3)+
#   scale_color_manual(values = c("#2F63A0","#7D9099"))+
#   geom_smooth(method = lm, color="#484848",lwd = 1, se=FALSE, formula =)
# 
# ggplot(CaseStudy2_data, aes(x=RelationshipSatisfaction,y = YearsAtCompany, color=EmployeeStatus )) +
#    geom_point(size=3)+
#   scale_color_manual(values = c("#2F63A0","#7D9099"))+
#   geom_smooth(method = lm, color="#484848",lwd = 1, se=FALSE)
```

####What are the top 3 factors that lead to attrition?

Using the Stepwise Variable selction method we have determined that the most effective predictors of years with the company are Years with Current Manager, Training Times Last Year,	Years In CurrentRole,	Years Since Last Promotion,	Number of Companies Worked,	Age,	Monthly Income,	Job Involvement, Percent Salary Hike and DailyRate.

These factors can be used to predict how long in years an employee will stay with the company using a statistical formula.

We advise caution be used in decision making based on the following variables for ethical or even legal reasons:  

* Gender  
* Marital Status  
* Relationship Satisfaction  
* Total Working Years
* Age

Our model actually only uses one of these factors (Age) which if used as a factor in decision making could be considered discriminatory therefore this analysis should be used with caution.

Inference can only be drawn to the employees in our dataset and not an larger external population.
```{r}
#Reference  https://cran.r-project.org/web/packages/olsrr/vignettes/variable_selection.html
#Create dataset of only the numeric categoricals
MyPredictionData<-data.frame(cbind(CaseStudy2_data$Age, CaseStudy2_data$DailyRate, CaseStudy2_data$DistanceFromHome, CaseStudy2_data$Education, CaseStudy2_data$EnvironmentSatisfaction, CaseStudy2_data$HourlyRate,CaseStudy2_data$JobInvolvement, CaseStudy2_data$JobLevel, CaseStudy2_data$JobSatisfaction, CaseStudy2_data$MonthlyIncome, CaseStudy2_data$MonthlyRate, CaseStudy2_data$NumCompaniesWorked, CaseStudy2_data$PercentSalaryHike, CaseStudy2_data$PerformanceRating, CaseStudy2_data$RelationshipSatisfaction,CaseStudy2_data$TrainingTimesLastYear, CaseStudy2_data$WorkLifeBalance,CaseStudy2_data$TotalWorkingYears,CaseStudy2_data$StockOptionLevel, CaseStudy2_data$YearsAtCompany, CaseStudy2_data$YearsInCurrentRole, CaseStudy2_data$YearsSinceLastPromotion, CaseStudy2_data$YearsWithCurrManager))
#Set the column Names of the MyPredictionData dataframe
colnames(MyPredictionData)<-c("Age","DailyRate","DistanceFromHome","Education","EnvironmentSatisfaction","HourlyRate","JobInvolvement","JobLevel","JobSatisfaction","MonthlyIncome","MonthlyRate","NumCompaniesWorked","PercentSalaryHike","PerformanceRating","RelationshipSatisfaction","StockOptionLevel","TotalWorkingYears","TrainingTimesLastYear","WorkLifeBalance","YearsAtCompany","YearsInCurrentRole","YearsSinceLastPromotion","YearsWithCurrManager")

#Other Methods of Variable Selection that were tested
  # Stepwise forward regression
  # model <- lm(YearsAtCompany ~ ., data = MyPredictionData)
  # ols_step_forward_p(model)
  # Stepwise backward regression
  # model <- lm(YearsAtCompany ~ ., data = MyPredictionData)
  # ols_step_backward_p(model)
  # stepwise aic regression
  # model <- lm(YearsAtCompany ~ ., data = MyPredictionData)
  # ols_step_both_aic(model)

# Stepwise regression
model <- lm(YearsAtCompany ~ ., data = MyPredictionData)
ols_step_both_p(model)
```
```{r, include=FALSE}
EmployeeRetention.lm<-lm(YearsAtCompany~YearsWithCurrManager+TrainingTimesLastYear+YearsInCurrentRole+YearsSinceLastPromotion+NumCompaniesWorked+Age+MonthlyIncome+JobInvolvement+PercentSalaryHike+DailyRate,data = MyPredictionData)
```
```{r}
#plot(EmployeeRetention.lm)
```
The top 3 factors that predict how long an employee will stay with the company in years are Years With Current Manager, Training Times Last Year and YearsInCurrentRole    

Though this analysis is significant it is merely a proof of concept; higher quality results can be achieved wiht additional time and resources for analyzing the data.

<<<<<<< HEAD
####Appendix
```{r}
ols_step_both_p(model)
EmployeeRetention.lm
plot(EmployeeRetention.lm)
```
=======
The business is also interested in learning about any job role specific trends that may exist in the data set (e.g., “Data Scientists have the highest job satisfaction”
>>>>>>> e3a6f392ef38102818de8e9aa562a311b0f1ce31
